### 2022-04-20 plos_latex_template.tex
Updates from Overleaf

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Ten simple rules to host an inclusive conference} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Rocío Joo\textsuperscript{1*\Yinyang}, 
Andrea Sánchez-Tapia\textsuperscript{2\Yinyang},
Sara Mortara\textsuperscript{3},
Yanina Bellini Saibene\textsuperscript{4,5},
Heather Turner\textsuperscript{6}, %\ddag},
Dorothea Hug Peter\textsuperscript{7}, %\ddag},
Natalia Soledad Morandeira\textsuperscript{8},
Matt Bannert\textsuperscript{9},
Batool Almazrouq\textsuperscript{10},
Elizabeth Hare\textsuperscript{11},
Laura Ación\textsuperscript{5,12},
Juan Pablo Narváez-Gómez\textsuperscript{13},
Marcela Alfaro Córdoba\textsuperscript{14},
Federico Marini\textsuperscript{15},
Rita Giordano\textsuperscript{16},
Silvia Canelón\textsuperscript{17},
Anicet Ebou\textsuperscript{18},
Adithi Upadhya\textsuperscript{19},
Joselyn Chávez\textsuperscript{20},
Janani Ravi\textsuperscript{21*}
%,3\textcurrency}
\\
\bigskip
\textbf{1} Global Fishing Watch, Washington, DC 20036, USA
\\
\textbf{2} Instituto de Pesquisas Jardim Botânico do Rio de Janeiro, Rio de Janeiro, Rio de Janeiro, Brazil
\\
\textbf{3} International Institute for Sustainability, Rio de Janeiro, Rio de Janeiro, Brazil
\\
\textbf{4} Instituto Nacional de Tecnología Agropecuaria, Ruta Nac. N° 5, Km. 580, Anguil, La Pampa, Argentina; Universidad Nacional Guillermo Brown, Buenos Aires, Argentina
\\
\textbf{5} MetaDocencia, Argentina
\\
\textbf{6} Department of Statistics, University of Warwick, Coventry, UK
\\
\textbf{7} Swiss Federal Institute for Forest, Snow and Landscape Research WSL, Birmensdorf, Switzerland
\\
\textbf{8} Instituto de Investigación e Ingeniería Ambiental, Universidad Nacional de San Martín - Consejo Nacional de Investigaciones Científicas y Técnicas, General San Martín, Buenos Aires, Argentina
\\
\textbf{9} KOF Swiss Economic Institute, ETH Zurich, Zurich, Switzerland
\\
\textbf{10} Institute of Systems, Molecular and Integrative Biology, University of Liverpool, Liverpool, L69 7ZB, UK; King Abdullah International Medical Research Center, Riyadh, Saudi Arabia
\\
\textbf{11} Dog Genetics LLC, Astoria, NY, USA
\\
\textbf{12} Instituto de Cálculo, FCEyN, UBA-CONICET, Buenos Aires, Argentina
\\
\textbf{13} Departamento de Botânica, Instituto de Biociências, Universidade de São Paulo, São Paulo, São Paulo, Brasil
\\
\textbf{14} Department of Statistics, University of California, Santa Cruz, CA, USA
\\
\textbf{15} Institute of Medical Biostatistics, Epidemiology and Informatics (IMBEI), University Medical Center Mainz, Mainz, Germany
\\
\textbf{16} Royal Society of Chemistry, Thomas Graham House (290), Science Park, Milton Road, Cambridge, UK
\\
\textbf{17} Department of Biostatistics, Epidemiology \& Informatics, University of Pennsylvania, Philadelphia, PA, USA.
\\
\textbf{18} Bioinformatics team, Département de Formation et de Recherches en Agriculture et Ressources Animales, Institut National Polytechnique Félix Houphouët-Boigny, BP 1313 Yamoussoukro, Côte d’Ivoire.
\\
\textbf{19} ILK Labs, Bengaluru, Karnataka, 560046, India
\\
\textbf{20} Instituto de Biotecnología, UNAM, Cuernavaca, Morelos, México
\\
\textbf{21} Departments of Pathobiology and Diagnostic Investigation, Microbiology and Molecular Genetics, Michigan State University, East Lansing, MI 48824, USA
\\
\bigskip

@@

Conferences are spaces to meet and to network within and across academic and technical fields, to learn about new advances, and share our work. They can help define career paths and create long-lasting collaborations and opportunities. 
However, these opportunities are not equal for all. 
This article introduces ten simple rules to host an inclusive conference based on the authors' recent experience organizing the 2021 edition of the useR! statistical computing conference, which attracted a broad range of participants from academia, industry, government, and the non-profit sector. 
Coming from different backgrounds, career stages, and even continents, we embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many people as possible.
The rules result from our lessons learned before, during, and after the organization of the conference. 
They have been written mainly for potential organizers and selection committees of conferences and contain multiple practical tips to help a variety of events become more accessible and inclusive. We see this as a starting point for conversations and efforts towards building more inclusive conferences across the world.
\hl{* Translated versions of the extended abstract are available in 10 languages in the Supplementary Information--Arabic, French, German, Italian, Japanese, Korean, Portuguese, Spanish, Tamil, and Thai.}

@@

% Roadmap of the paper
Here, we present a set of ten rules based on the lessons we learned before, during, and after the organization of the useR! conference.
\linelabel{methods-a} \hl{The rules were first drafted by the core team and members of the diversity, accessibility, and inclusion team of useR! 2021 who wrote their own ten simple rules (see Supplementary Information) based on the collective experience of organizing the conference and fueled by personal experiences and literature review.
A final list was proposed and then discussed and reviewed by the rest of the coauthors.} \linelabel{methods-b}
The rules are organized in three sections: three foundation rules, six design rules, and a continuity rule (\textbf{Fig. \ref{fig:diagram}}).
The \textbf{foundation rules} comprise key elements to conceive the work on diversity and inclusion in any conference. 
\textbf{Rule 1} is about setting a vision of diversity and inclusion that should guide all the efforts and decision making in the organization.
\textbf{Rule 2} focuses on how to create a safe and welcoming environment for all the attendees. 
\textbf{Rule 3} highlights the importance of starting with an inclusive and diverse organizing team and provides tips on work dynamics.
The \textbf{design rules} focus on weaving inclusion into the conference design process.
In \textbf{Rule 4}, we introduce multiple ways to counteract bias in the conference program (keynotes, program committee, abstract selection, and thematic sessions). 
\textbf{Rule 5} provides advice for designing an inclusive online component in virtual and hybrid conferences.
\textbf{Rule 6} focuses on accessibility practices to include people with disabilities.
In \textbf{Rule 7}, we provide suggestions to account for linguistic diversity. 
\textbf{Rule 8} offers tips for developing an inclusive communication strategy. 
In \textbf{Rule 9}, we address budgeting for inclusive practices and helping participants with affordable registration costs, scholarships, and other forms of financial support.
Finally, 
\textbf{Rule 10}, the \textbf{continuity} rule, emphasizes the importance of self-assessment and advocates for making the conference part of a long-term commitment to inclusion and for passing the torch to future organizers.

@@

The CoC response team must receive training on how to receive reports, respond to incidents, and communicate their responses to the whole community. The team should be prepared early during conference planning because CoC violations can happen before the conference begins (e.g.\ on social media, or in organization meetings).  
Responding to CoC reports can be one of the most emotionally intense tasks of the team. Resources permitting, consider adding funds for their work to your budget (see \textbf{Rule 9}).
We strongly recommend reading `How to Respond to Code of Conduct Reports' \cite{auroraHowRespondCode2019} as an excellent starting point for building a strong CoC response team.

@@

\linelabel{subteam-1} \hl{Create sub-teams to distribute fairly the large amount of work that organizing a conference represents. 
Having a dedicated DEI team may be an excellent way to give the topic the attention and energy it requires.
The core team should support the DEI initiatives and advocate for inclusion across all conference areas.} \linelabel{subteam-2} 
It is important, though, to ensure that organizers with diverse backgrounds are not restricted to only work on diversity and inclusion aspects; 
every member should have the freedom to choose which areas of the conference they want to contribute to.
When it comes to team leadership, do not expect self-nomination and voting to work as infallible mechanisms to fill these positions equitably. 
Instead, nominate folks directly and offer them leadership positions in the organization team, especially those positions that would be typically occupied by people from privileged groups.
Bear in mind that some people may lack the institutional or financial support to put time and effort into the organization tasks and do not have the luxury to commit for free; consider reallocating resources to give them a stipend for their work (see \textbf{Rule 9}).
In any case, try to find meaningful ways to acknowledge everybody's work. Use the conference spaces (e.g. webpage, opening/closing ceremony at the conference, social media) to give proper credit to the people who are making the event possible.
And most importantly, take care of your team and their well-being! Check on them regularly and make sure that everyone is comfortable. This represents a great amount of work and you will need to support each other for the long haul.

@@

The conference program may lack diversity in the invited roles––-such as keynote speakers, program/scientific committee, session chairs, panelists––-, the accepted abstracts, and even the scope of the topics covered. 
It is important to counteract biases in all of these aspects.
When choosing or inviting people for visible and valued roles in the conference, it is likely that there will not be much diversity in the first pool of names \cite{dwyerNoticeWhoScience2021,swartzScienceValueDiversity2019,wongBuildDiversityScience2020,dignazioUnicornsJanitorsNinjas2020}. \linelabel{invited}
Rather than deterring you, this should encourage you to go beyond your current networks to look for, reach out to, invite, encourage, and on-board great people that are not routinely in the spotlight, making sure that these roles do not reinforce existing privileges.

Biases in the final list of abstracts begin with self-selection: some potential conference participants may not feel confident enough to submit their work, especially to large and prestigious conferences.
To mitigate such self-selection, some communities have pre-submission mechanisms to aid their members prepare and receive feedback for their abstract (e.g. the Latin-R and R-Ladies communities have volunteers to help their members improve their conference abstracts).
If this does not exist in your field yet, consider asking members of your community to provide a similar kind of mentorship.
\linelabel{multilingual-submission}\hl{To mitigate self-selection from non-native speakers, consider accepting submissions in several languages} (see \textbf{Rule 7}).

The abstract review process can be subjected to implicit bias \cite{ross_everyday_2020} when folks in the selection committee unconsciously assign a positive or negative value to the names and affiliations of authors—and their perceived origin, ethnicity, gender, or native language—, affecting the chances to get their work accepted.\linelabel{bias} 
Selection committees need to be reminded of such biases in the evaluation process, to try to avoid judging more scrupulously the work of people perceived as part of minoritized groups or being kinder when reviewing the abstracts from those perceived as part of privileged groups and considering their work as more relevant \cite{swartzScienceValueDiversity2019, andersson_implicit_2019}.
Evaluate the plausibility of anonymizing \hl{the submission information as much as possible} \linelabel{anonymous}, and
define the most appropriate strategy for reviewing (e.g. double blind, double open, or single blind reviews; see \cite{numfocus_discover_2021}).
\linelabel{criteria-1} \hl{Even well-meaning and carefully crafted processes can be prone to reproduce the biases present in academia and society. It is, therefore, essential to examine the final list of selected abstracts to check if there is a lack of diversity. If proposals with similar quality were still judged differently because of unconscious bias, consider giving preference to the work from people in minoritized groups.
This manual examination at the final stage is unlikely to be reproducible. Nonetheless, the whole process should be as transparent as possible: evaluation criteria should be clearly stated and shared with the reviewers and authors, preferably during the call for abstracts.} \linelabel{criteria-2}
\linelabel{video-1}\hl{It is worth noting that self-selection and unconscious bias may be aggravated by the use of video abstracts as an alternative to written ones, affecting people without the resources to create good quality videos, people with disabilities, non-native speakers, racialized people, and people with diverse body types, among others.} \cite{spitschanVideoGrantProposals2021}\linelabel{video-2}

@@

If you are organizing a hybrid event, decide on the relative importance that online and in-person components will take \cite{bajpai_towards_2021}. The most inclusive practice would be to give equal weight to both components, articulating them well and maximizing the experience of everyone attending the conference.
The audience, chairs, and presenters should all be able to interact regardless of their in-person or remote status. 
To best integrate in-person and remote attendees, all questions and comments should pass through a single online system. 
The physical venue would have to include multiple ways to connect with online folks such as cameras and microphones to allow them to follow who is speaking at the in-person space, and tablets to help in-person attendees interact with remote participants.

@@

% Rocío: main idea: help take the burden out of participants for more inclusive participation
When determining the registration rates, the socioeconomic context of participants, their career status, and their country of origin—if applicable—should be taken into account  \cite{sarabipourChangingScientificMeetings2021, andalibPostdocQueueLabour2018, kaplanPostdocNot2012}
(see \cite{canelon2021cost} for an example of conversion rates based on country of origin and career status).\linelabel{example-countries}
Resources permitting, enable a `pay what you can' option. You could also aim to have a conference with no registration costs; this could especially apply for online events, but bear in mind that free events have a lower attendance rate than paid events \cite{eventbrite_ultimate_2017}. 
Scholarships to attend the in-person component of the conference are an additional way to boost participation of people from minoritized groups, by offering support for travel and lodging expenses.
Design the diversity-related scholarship to be explicit about the groups you want to support, and be transparent about the criteria for evaluation to avoid self-selection (refraining from applying). 
To assign these scholarships, conferences usually ask for cover letters or applications, which can be time-consuming and emotionally demanding. 
Simplifying the process of requesting financial support, especially for 
people who lack the time and resources (e.g. because of family responsibilities) could greatly increase their chances of applying for the support. 
Additional aid for attendees could also be offered—e.g. paying for child care or internet connection services for the online component. 
Bear in mind that transferring money internationally could be a cumbersome administrative process and significantly augment the burden of already marginalized groups. Whenever possible, facilitate alternate ways for money transfer (e.g. book flights, hotel reservations, waive conference fees, give gift cards).

@@

% Rocío: main idea: assess how inclusive was the conference
\linelabel{account-3}As conference organizers, one of the key post-conference tasks is to make a final assessment of the achievement of the diversity and inclusion goals. 
You can rely on the diversity goals defined at the beginning of the planning process (\textbf{Rule 1}). 
\linelabel{info}\hl{The organizers will likely be able} to evaluate the outcomes associated with these goals directly from the information collected from the organizing team, speakers, and participants (e.g.\ \hl{city/state/country of origin} or preferred language).
The planning committee may have set goals concerning how welcomed and included participants felt during the conference, and need to assess if the implemented practices translated into true inclusion. 
Surveys, focus groups, and informal conversations during and after the conference are good ways to collect this information and great opportunities to receive feedback from the attendees. \linelabel{info2}

% Pass the information
The feedback received from the participants and the whole assessment of the conference should be documented and shared with \linelabel{data-share} \hl{those who helped define the inclusion goals, the attendees, and} future organizers. 
\hl{Ensure that the participants' privacy and anonymity are respected; ask for consent and consider binning/grouping the data into categories when there is a low number of respondents}.\linelabel{data-share2}
Gather and do your best to organize all the documentation produced during the organization process---or even after. Timelines, checklists, budgeting tips, email templates, accessibility checks and guidelines, are a few of many resources that can be shared with your conference successors. 
Create a list with contact information of the people of the organizing team, as well as previous and potential speakers---if they agree to share their contact information---, and pass it along.
For conference series, it could be a good idea to create an official document (e.g. a wiki, knowledge base or similar guide, see for instance \cite{sanchez-tapia_user_2021-2}) that can be used and kept up to date by future organizers.
If you are part of a stable meeting committee (overseeing multiple editions), encourage organizers to follow these Rules and set up new standards for inclusion: be explicit about the inclusive spirit of the conference and the practices that future organizers should commit to when publishing the call for conference organizers and defining the selection criteria.

@@

HT was supported by the Engineering and Physical Sciences Research Council [EPSRC EP/V052128/1] during the preparation of this article and by The R Foundation for Statistical Computing during the organization of useR! 2021. HT is a board member of the The R Foundation for Statistical Computing, serving on the R Foundation conference committee that arranges hosts for the useR! conference series and chairing Forwards, the R Foundation taskforce for women and under-represented groups. The R Foundation is currently funding YBS to support the useR! 2022 organization team. HT was administrator and volunteer for the R Project Google Season of Docs 2021 project, that funded work by AST (writer) and YBS (steering committee).
The organization of useR! 2021 was largely a volunteer effort, but stipends were available from the conference funds for specific roles. RJ, AST, SM, BA, NM, AU, and AE received code of conduct team stipends; AST, YBS, JPNG, MAC, RG, and JC received zoom host stipends; AST and MAC received stipends for their unrelenting efforts as diversity and social program leads, respectively.

For the purpose of open access, the authors have applied a Creative Commons Attribution (CC BY) license to any Author Accepted Manuscript version arising from this submission.

>>>>

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Ten simple rules to host an inclusive conference} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Rocío Joo\textsuperscript{1*\Yinyang}, 
Andrea Sánchez-Tapia\textsuperscript{2\Yinyang},
Sara Mortara\textsuperscript{3},
Yanina Bellini Saibene\textsuperscript{4,5},
Heather Turner\textsuperscript{6}, %\ddag},
Dorothea Hug Peter\textsuperscript{7}, %\ddag},
Natalia Soledad Morandeira\textsuperscript{8},
Matt Bannert\textsuperscript{9},
Batool Almazrouq\textsuperscript{10},
Elizabeth Hare\textsuperscript{11},
Laura Ación\textsuperscript{5,12},
Juan Pablo Narváez-Gómez\textsuperscript{13},
Marcela Alfaro Córdoba\textsuperscript{14},
Federico Marini\textsuperscript{15},
Rita Giordano\textsuperscript{16},
Silvia Canelón\textsuperscript{17},
Anicet Ebou\textsuperscript{18},
Adithi Upadhya\textsuperscript{19},
Joselyn Chávez\textsuperscript{20},
Janani Ravi\textsuperscript{21*}
%,3\textcurrency}
\\
\bigskip
\textbf{1} Global Fishing Watch, Washington, DC 20036, USA
\\
\textbf{2} Instituto de Pesquisas Jardim Botânico do Rio de Janeiro, Rio de Janeiro, Rio de Janeiro, Brazil
\\
\textbf{3} International Institute for Sustainability, Rio de Janeiro, Rio de Janeiro, Brazil
\\
\textbf{4} Instituto Nacional de Tecnología Agropecuaria, Ruta Nac. N° 5, Km. 580, Anguil, La Pampa, Argentina; Universidad Nacional Guillermo Brown, Buenos Aires, Argentina
\\
\textbf{5} MetaDocencia, Argentina
\\
\textbf{6} Department of Statistics, University of Warwick, Coventry, UK; The R Foundation
\\
\textbf{7} Swiss Federal Institute for Forest, Snow and Landscape Research WSL, Birmensdorf, Switzerland
\\
\textbf{8} Instituto de Investigación e Ingeniería Ambiental, Universidad Nacional de San Martín - Consejo Nacional de Investigaciones Científicas y Técnicas, General San Martín, Buenos Aires, Argentina
\\
\textbf{9} KOF Swiss Economic Institute, ETH Zurich, Zurich, Switzerland
\\
\textbf{10} Institute of Systems, Molecular and Integrative Biology, University of Liverpool, Liverpool, L69 7ZB, UK; King Abdullah International Medical Research Center, Riyadh, Saudi Arabia
\\
\textbf{11} Dog Genetics LLC, Astoria, NY, USA
\\
\textbf{12} Instituto de Cálculo, FCEyN, UBA-CONICET, Buenos Aires, Argentina
\\
\textbf{13} Departamento de Botânica, Instituto de Biociências, Universidade de São Paulo, São Paulo, São Paulo, Brasil
\\
\textbf{14} Department of Statistics, University of California, Santa Cruz, CA, USA
\\
\textbf{15} Institute of Medical Biostatistics, Epidemiology and Informatics (IMBEI), University Medical Center Mainz, Mainz, Germany
\\
\textbf{16} Royal Society of Chemistry, Thomas Graham House (290), Science Park, Milton Road, Cambridge, UK
\\
\textbf{17} Department of Biostatistics, Epidemiology \& Informatics, University of Pennsylvania, Philadelphia, PA, USA.
\\
\textbf{18} Bioinformatics team, Département de Formation et de Recherches en Agriculture et Ressources Animales, Institut National Polytechnique Félix Houphouët-Boigny, BP 1313 Yamoussoukro, Côte d’Ivoire.
\\
\textbf{19} ILK Labs, Bengaluru, Karnataka, 560046, India
\\
\textbf{20} Instituto de Biotecnología, UNAM, Cuernavaca, Morelos, México
\\
\textbf{21} Departments of Pathobiology and Diagnostic Investigation, Microbiology and Molecular Genetics, Michigan State University, East Lansing, MI 48824, USA
\\
\bigskip

@@

Conferences are spaces to meet and to network within and across academic and technical fields, to learn about new advances, and share our work. They can help define career paths and create long-lasting collaborations and opportunities. 
However, these opportunities are not equal for all. 
This article introduces ten simple rules to host an inclusive conference based on the authors' recent experience organizing the 2021 edition of the useR! statistical computing conference, which attracted a broad range of participants from academia, industry, government, and the non-profit sector. 
Coming from different backgrounds, career stages, and even continents, we embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many people as possible.
The rules result from our lessons learned before, during, and after the organization of the conference. 
They have been written mainly for potential organizers and selection committees of conferences and contain multiple practical tips to help a variety of events become more accessible and inclusive. We see this as a starting point for conversations and efforts towards building more inclusive conferences across the world.
\hl{* Translated versions of the extended abstract are available in ten languages in the Supplementary Information: Arabic, French, German, Italian, Japanese, Korean, Portuguese, Spanish, Tamil, and Thai.}

@@

% Roadmap of the paper
Here, we present a set of ten rules based on the lessons we learned before, during, and after the organization of the useR! conference.
\linelabel{methods-a} \hl{The rules were first drafted by the core team and the members of the diversity, accessibility, and inclusion team of useR! 2021 who wrote their own ten simple rules (see Supplementary Information) based on the collective experience of organizing the conference and fueled by personal experiences and literature review.
A final list was proposed and then discussed and reviewed by the rest of the coauthors.} \linelabel{methods-b}
The rules are organized in three sections: three foundation rules, six design rules, and a continuity rule (\textbf{Fig. \ref{fig:diagram}}).
The \textbf{foundation rules} comprise key elements to conceive the work on diversity and inclusion in any conference. 
\textbf{Rule 1} is about setting a vision of diversity and inclusion that should guide all the efforts and decision making in the organization.
\textbf{Rule 2} focuses on how to create a safe and welcoming environment for all the attendees. 
\textbf{Rule 3} highlights the importance of starting with an inclusive and diverse organizing team and provides tips on work dynamics.
The \textbf{design rules} focus on weaving inclusion into the conference design process.
In \textbf{Rule 4}, we introduce multiple ways to counteract bias in the conference program (keynotes, program committee, abstract selection, and thematic sessions). 
\textbf{Rule 5} provides advice for designing an inclusive online component in virtual and hybrid conferences.
\textbf{Rule 6} focuses on accessibility practices to include people with disabilities.
In \textbf{Rule 7}, we provide suggestions to account for linguistic diversity. 
\textbf{Rule 8} offers tips for developing an inclusive communication strategy. 
In \textbf{Rule 9}, we address budgeting for inclusive practices and helping participants with affordable registration costs, scholarships, and other forms of financial support.
Finally, 
\textbf{Rule 10}, the \textbf{continuity} rule, emphasizes the importance of self-assessment and advocates for making the conference part of a long-term commitment to inclusion and for passing the torch to future organizers.

@@

The CoC response team must receive training on how to receive reports, respond to incidents, and communicate their responses to the whole community. The team should be prepared early during conference planning because CoC violations can happen before the conference begins (e.g.\ on social media or in organization meetings).  
Responding to CoC reports can be one of the most emotionally intense tasks of the team. Resources permitting, consider adding funds for their work to your budget (see \textbf{Rule 9}).
We strongly recommend reading `How to Respond to Code of Conduct Reports' \cite{auroraHowRespondCode2019} as an excellent starting point for building a strong CoC response team.

@@

\linelabel{subteam-1} \hl{Create sub-teams to distribute fairly the large amount of work that organizing a conference represents. 
Having a dedicated DEI team may be an excellent way to give the topic the attention and energy it requires.
The core team should support the DEI initiatives and advocate for inclusion across all conference areas.} \linelabel{subteam-2} 
It is important, though, to ensure that organizers with diverse backgrounds are not restricted to only work on diversity and inclusion aspects; 
every \hl{person} should have the freedom to choose which areas of the conference they want to contribute to.
When it comes to team leadership, do not expect self-nomination and voting to work as infallible mechanisms to fill these positions equitably. 
Instead, nominate folks directly and offer them leadership positions in the organizing team, especially those positions that would be typically occupied by people from privileged groups.
\hl{Moreover,} some people may lack the institutional or financial support to put time and effort into the organization tasks and do not have the luxury to commit for free; consider reallocating resources to give them a stipend for their work (see \textbf{Rule 9}).
In any case, try to find meaningful ways to acknowledge everybody's work. Use the conference spaces (e.g. webpage, opening/closing ceremony at the conference, social media) to give proper credit to the people who are making the event possible.
And most importantly, take care of your team and their well-being! Check on them regularly and make sure that everyone is comfortable. This represents a great amount of work and you will need to support each other for the long haul.

@@

\linelabel{invited}The conference program may lack diversity in the invited roles––-such as keynote speakers, program/scientific committee, session chairs, panelists––-, the accepted abstracts, and even the scope of the topics covered. 
It is important to counteract biases in all of these aspects.
When choosing or inviting people for visible and valued roles in the conference, it is likely that there will not be much diversity in the first pool of names \cite{dwyerNoticeWhoScience2021,swartzScienceValueDiversity2019,wongBuildDiversityScience2020,dignazioUnicornsJanitorsNinjas2020}.
Rather than deterring you, this should encourage you to go beyond your current networks to look for, reach out to, invite, encourage, and on-board great people that are not routinely in the spotlight, making sure that these roles do not reinforce existing privileges.\linelabel{invited-2}

Biases in the final list of abstracts begin with self-selection: some potential conference participants may not feel confident enough to submit their work, especially to large and prestigious conferences.
To mitigate such self-selection, some communities have pre-submission mechanisms to aid their members prepare and receive feedback for their abstract (e.g. the Latin-R and R-Ladies communities have volunteers to help their members improve their conference abstracts).
If this does not exist in your field yet, consider asking members of your community to provide a similar kind of mentorship.
\linelabel{multilingual-submission}\hl{To mitigate self-selection from non-native speakers, consider accepting submissions in several languages} (see \textbf{Rule 7}).\linelabel{multilingual-submission2}

The abstract review process can be subjected to implicit bias \cite{ross_everyday_2020} when folks in the selection committee unconsciously assign a positive or negative value to the names and affiliations of authors—and \linelabel{bias} their perceived origin, ethnicity, gender, or native language—, affecting the chances to get their work accepted.
Selection committees need to be reminded of such biases in the evaluation process, to try to avoid judging more scrupulously the work of people perceived as part of minoritized groups or being kinder when reviewing the abstracts from those perceived as part of privileged groups and considering their work as more relevant \cite{swartzScienceValueDiversity2019, andersson_implicit_2019}.
Evaluate the plausibility of anonymizing \hl{the submission information as much as possible}\linelabel{anonymous} and
define the most appropriate strategy for reviewing (e.g. double blind, double open, or single blind reviews; see \cite{numfocus_discover_2021}).
\linelabel{criteria-1} \hl{Even well-meaning and carefully crafted processes can be prone to reproduce the biases present in academia and society. It is, therefore, essential to examine the final list of selected abstracts to check if there is a lack of diversity. If proposals with similar quality were still judged differently because of unconscious bias, consider giving preference to the work from people in minoritized groups.
This manual examination at the final stage is unlikely to be reproducible. Nonetheless, the whole process should be as transparent as possible: evaluation criteria should be clearly stated and shared with the reviewers and authors, preferably during the call for abstracts.} \linelabel{criteria-2}
\linelabel{video-1}\hl{It is worth noting that self-selection and unconscious bias may be aggravated by the use of video abstracts as an alternative to written ones, affecting people without the resources to create good quality videos, people with disabilities, non-native speakers, racialized people, and people with diverse body types, among others.} \cite{spitschanVideoGrantProposals2021}\linelabel{video-2}

@@

If you are organizing a hybrid event, decide on the relative importance that online and in-person components will take \cite{bajpai_towards_2021}. The most inclusive practice would be to give equal weight to both components, articulating them well and maximizing the experience of everyone attending the conference.
The audience, chairs, and presenters should all be able to interact regardless of their in-person or remote status. 
To best integrate in-person and remote attendees, all questions and comments should pass through a single online system. 
The physical venue would have to include multiple ways to connect with online folks such as cameras and microphones to allow them to follow who is speaking at the in-person space, and tablets to help in-person attendees \hl{without suitable personal devices to interact with remote participants during Q\&A or networking sessions}.\linelabel{tablets}

@@

% Rocío: main idea: help take the burden out of participants for more inclusive participation
When determining the registration rates, the socioeconomic context of participants, their career status, and their country of origin—if applicable—should be taken into account  \cite{sarabipourChangingScientificMeetings2021, andalibPostdocQueueLabour2018, kaplanPostdocNot2012}
(see \cite{canelon2021cost} for an example of conversion rates based on country of origin and career status).\linelabel{example-countries}
Resources permitting, enable a `pay what you can' option. You could also aim to have a conference with no registration costs; this could especially apply for online events, but bear in mind that free events have a lower attendance rate than paid events \cite{eventbrite_ultimate_2017}. 
Scholarships to attend the in-person component of the conference are an additional way to boost participation of people from minoritized groups by offering support for travel and lodging expenses.
Design the diversity-related scholarship to be explicit about the groups you want to support and be transparent about the criteria for evaluation to avoid self-selection (refraining from applying). 
To assign these scholarships, conferences usually ask for cover letters or applications, which can be time-consuming and emotionally demanding. 
Simplifying the process of requesting financial support, especially for 
people who lack the time and resources (e.g. because of family responsibilities) could greatly increase their chances of applying for the support. 
Additional aid for attendees could also be offered—e.g. paying for child care or internet connection services for the online component. 
Bear in mind that transferring money internationally could be a cumbersome administrative process and significantly augment the burden of already marginalized groups. Whenever possible, facilitate alternate ways for money transfer (e.g. book flights, hotel reservations, waive conference fees, give gift cards).

@@

% Rocío: main idea: assess how inclusive was the conference
\linelabel{account-3}As conference organizers, one of the key post-conference tasks is to make a final assessment of the achievement of the diversity and inclusion goals. 
You can rely on the diversity goals defined at the beginning of the planning process (\textbf{Rule 1}). 
\linelabel{info}\hl{You will likely be able} to evaluate the outcomes associated with these goals directly from the information collected from the organizing team, speakers, and participants (e.g.\ \hl{city/state/country of origin} or preferred language).
You may have set goals concerning how welcomed and included participants felt during the conference, and need to assess if the implemented practices translated into true inclusion. 
Surveys, focus groups, and informal conversations during and after the conference are good ways to collect this information and great opportunities to receive feedback from the attendees. \linelabel{info2}

% Pass the information
\linelabel{data-share}The feedback received from the participants and the whole assessment of the conference should be documented and shared with \hl{those who helped define the inclusion goals, the attendees, and} future organizers. 
\hl{Ensure that the participants' privacy and anonymity are respected; ask for consent and consider binning/grouping the data into categories when there is a low number of respondents}.\linelabel{data-share2}
Gather and do your best to organize all the documentation produced during the organization process---or even after. Timelines, checklists, budgeting tips, email templates, accessibility checks and guidelines, are a few of many resources that can be shared with your conference successors. 
Create a list with contact information of the members of the organizing team, as well as previous and potential speakers---if they agree to share their contact information---, and pass it along.
For conference series, it could be a good idea to create an official document (e.g. a wiki, knowledge base or similar guide, see for instance \cite{sanchez-tapia_user_2021-2}) that can be used and kept up to date by future organizers.
If you are part of a stable meeting committee (overseeing multiple editions), encourage organizers to follow these Rules and set up new standards for inclusion: be explicit about the inclusive spirit of the conference and the practices that future organizers should commit to when publishing the call for conference organizers and defining the selection criteria.

@@

HT was supported by the Engineering and Physical Sciences Research Council [EPSRC EP/V052128/1] during the preparation of this article and by The R Foundation for Statistical Computing for the organization of useR! 2021. HT is a board member of The R Foundation for Statistical Computing, serving on the R Foundation conference committee that arranges hosts for the useR! conference series and chairing Forwards, the R Foundation taskforce for women and under-represented groups. The R Foundation is currently funding YBS to support the useR! 2022 organization team. HT was administrator and volunteer for the R Project Google Season of Docs 2021 project, that funded work by AST (writer) and YBS (steering committee).
The organization of useR! 2021 was largely a volunteer effort, but stipends were available from the conference funds for specific roles. RJ, AST, SM, BA, NM, AU, and AE received code of conduct team stipends; AST, YBS, JPNG, MAC, RG, and JC received zoom host stipends; AST and MAC received stipends for their unrelenting efforts as diversity and social program leads, respectively. JR is funded by start-up award from Michigan State University; publication cost will waived due to PLoS-MSU partnership.

\section*{Data availability}
For the purpose of open access, the authors have applied a Creative Commons Attribution (CC BY) license to any Author Accepted Manuscript version arising from this submission. The manuscript and the supporting docs are available at: https://github.com/rociojoo/Ten-simple-rules-inclusive-conference.
